### 2021-04-20 DRAFT.Rmd
mana edits

---
title: "ST563 Final Project"
author: "Final Project Group 6"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    toc_depth: 1
fontsize: 12pt
---

```{r setup, include=FALSE}
# Set default chunk options
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE, fig.align = "center")

# Load necessary libraries
library(tidyverse) # so many things
library(knitr) # making nice tables
library(caret) # training models & CV
library(leaps) # subset selection
library(corrplot) # visualize correlation
library(rmarkdown) # for output documents
library(rpart) # fitting classification tree
library(rpart.plot) # visualization of classification tree
library(randomForest) # fitting random forest model
library(kableExtra) # making bolded headers for tables
library(gridExtra) # stitching graphs together
library(tree)

# Read in data set
wine <- read.csv2(file = "winequality-red.csv", header = TRUE, dec = ".",
                  colClasses = c(rep("double", 7), rep("double", 4), "integer"))

# Set seed for reproducibility
set.seed(702)
```

\newpage

# Introduction

There has been a push towards the implementation of different analytics solutions within the agriculture industry, especially in the realms of crop health and yield. These sort of analyses can help farmers reduce waste and improve profits. In particular, the viticulture industry stands to benefit from such methods, being that the production of wine is lengthy and multifaceted. Winemakers can work to optimize revenue by analyzing their input and output (crops and wines). Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality. The question this study is most interested in is to determine if there is a way to predict the quality of a wine without having to use taste testers. The team is hypothesizing that both the classifation and regression methods can accurately describe the quality of wine.

In this study, the objective is to model and predict the quality of wine using a variety of statistical learning techniques. This will be done by performing both regression and classification methods, with regression ranking the quality of the wines on a scale of zero to ten, with zero being the lowest quality and ten being the highest quality. The classification models will be classifying the wines into two categories, good and bad, with wines in the bad category having a score of zero to five, and wines in the good category having a quality score between six and ten. The assessment of the prediction accuracy will be performed by the following methods:

```{r methods}
# Make table of methods
cl <- c("Classification Tree", "Random Forest")
reg <- c("Multiple Linear Regression (MLR)", "Random Forest")
kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods") %>% 
  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% 
  kable_styling(latex_options = "hold_position")
```

# Required Libraries
To run the code for the project, the following libraries are required:

  * `caret`: to do the heavy lifting of training and tuning the models
  * `tidyverse`: for all the data reading and wrangling
  * `knitr`: for rmarkdown table outputs
  * `rmarkdown`: for output documents
  * `corrplot`: for correlation structure visualization
  * `leaps`: for subset selection
  * `rpart`: fitting the basic classification tree
  * `rpart.plot`: visualization of the classification tree
  * `randomForest`: fitting random forest models

# Data
The data set used is the Wine Quality dataset from the UCI Machine Learning Repository. The data set is comprised of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine. For each wine, various physical and chemical features of each wine was measured, including a measurement of quality given by an expert wine taster. The objective is to use the different features to predict the quality of the wine using different machine learning techniques.

## Variable Descriptions  
Each entry in the dataset represents the different metrics of the following attributes of a single type of wine.
  * **fixed acidity**: the quantity of fixed acids found in the wines. The predominant fixed acids found in wines are tartaric, malic, citric, and succinic, all of which come from the grapes except for the succinic acid, which comes from the yeast in fermentation. 
  * **volatile acidity**: the steam distillable acids present in wine, primarily acetic acid but also lactic, formic, butyric, and propionic acids. These acids generally come from the fermentation process.
  * **citric acid**: added to the wine as a natural preservative or for acidity and tartness.
  * **residual sugar**: the quantity of sugar left in the wine after the fermentation process.
  * **chlorides**: the amount of salt in a wine.
  * **free sulfur dioxide**: the amount of $SO_2$ that is not bound to other molecules. 
  * **total sulfur dioxide**: total amount of $SO_2$ in the wine. Sulfur Dioxide is used throughout all stages of the winemaking process to prevent oxidation and bacteria growth.
  * **density**: density of the wine.
  * **pH**: pH of the wine.
  * **sulphates**: quantity of sulphates in the wine. Sulphates are used as a preservative.
  * **alcohol**: alcohol content by volume.
  * **quality**: score of quality of the wine given by expert tasters (score between 0 and 10).

## Data Exploration
The first thing that was done to the data upon reading it was to use the numeric variables and create pairwise scatterplots. This way, one can obtain a greater understanding of how the variables relate to one another, as well as if there is any correlation structure and what it looks like. If the variables are related, there may be problems moving forward, so it is important to know these relationships before the analysis is performed.

```{r pairs, out.width='4in', out.height='3.5in'}
# Create pairwise scatterplots
pairs(wine[,c(2,5,7,8,9,10,11)])
```
Based on the pairwise scatterplots, one can conclude that pH may be linearly correlated with density, sulphates, and alcohol. This leads us to believe that there is some relationship between the chemical properties of the wines. here also appears to be a relationship between density and alcohol. The following correlation plot gives a numeric summary of the linear relationship between the variables.

```{r corrplot, echo = FALSE, out.height='3.5in'}
# Create visualization of correlation structure
p.mat <- cor.mtest(wine[,c(2,5,7,8,9,10,11)])$p
corrplot(cor(wine[,c(2,5,7,8,9,10,11)]),
         method = "color", type = "upper", order = "hclust", 
         number.cex = .7, addCoef.col = "black",
         tl.col = "black", tl.srt = 90,
         p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         diag = T)
```

The only linear correlations that are notably high are between alcohol and density. This makes sense because the density of alcohol is slightly less than water, meaning that as the concentration of alcohol gets higher, the density will naturally decrease. Similarly, the sulphates and chlorides seem to have a slightly positive linear relationship, while pH and density have a slightly negative linear relationship.

### Good Wine Qualities v. Bad Wine Qualities

After a split in the data was performed to pull out the wines classified as good and the wines classified as bad. Further analysis was performed on each group to see if any of the variables were significantly different between the two groups.

```{r good v bad}
good.wine <- subset(wine, quality > 5)
bad.wine <- subset(wine, quality <= 5)
# summary(good.wine)
# summary(bad.wine)
```
The two groups were similar in most of the variables. Two notable differences are that the wines classified as "good" had a lower mean value than "bad" wines for total sulfur dioxide. Good wines had a mean of 39, which was significantly lower than the mean of the "bad" wines at 54.6. Another difference to note is the difference in the mean for alcohol. The good wines had a mean alcohol content at 10.8, while the bad wines had a lower mean alcohol content around 9.9. There were also differences in fixed acidity, with the bad wines having a lower mean. There were also several variables that don't seem to change between the two groups including pH, residual sugar, chlorides, and density.

## Data Processing
```{r partition}
# Prepare data for classification
wine$class <- ifelse(wine$quality <= 5, "low", "high") %>% as.factor()

# Create Test and Training Sets
ind <- sample(length(wine$class), 0.8*length(wine$class))

# Create train/test data for H1
trn <- wine[ind,]
tst <- wine[-ind,]
```

Fortunately, the dataset was very tidy; there were no missing values. The data was first split into a training and test data set. The training data set uses 80% of the observations and the test set contains the remaining 20%. The models are built on the training set and evaluated on the test set. This is important so that the models are not overfit. A variable was then created for classification, transforming the quality variable into a binary "low" or "high" response as mentioned in the introduction.

# Classification Models
## Classification Tree

THIS SECTION NEEDS TO BE EDITED

The basic classification tree is based on partitioning the data into subgroups using simple binary splitting. Initially, all objects are considered a single group. Then, the group is repeatedly split into two subgroups based on the criteria of a certain variable. In the classification setting, an observation is classified in a specific region with majority vote. Trees are grown as big as possible and then pruned using cost-complexity pruning. This is done to ensure the data is not being overfit. The tradeoff here is that pruning increases the bias. The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods. Additionally, the classification tree is not very computationally expensive, unlike the random forest model.

```{r tree, out.height='3in'}
# Set seed for reprodicibility
set.seed(702)

# 1. full size decision tree
tree.class <- tree(class ~ . - quality, trn)
plot(tree.class); text(tree.class, pretty = 0)
tree.class.pred <- predict(tree.class, tst, type = 'class')
tree.class.train_err <- mean(trn$class != predict(tree.class, type = 'class'))
tree.class.test_err <- mean(tst$class != tree.class.pred)
# tree.class.train_err
# tree.class.test_err
```

The resulting classification tree creates splits on 4 variables: Alcohol, Total Sulfur Dioxide, Sulphates, and Volatile Acidity. The tree has a test error rate of `r round(tree.class.test_err, 3)`. This test rate is fairly low, but as mentioned before, if one conducts pruning on the tree, the error rate can decrease and a more accurate model is created.

Pruning is performed on the tree using cross validation, the tree is built many times with a different number of splits and the tree that produces the smallest error rate is selected. Cross validation on this model shows that the best model is the model built with 8 splits. The resulting pruned model will now have 8 terminal nodes and still uses the four variables in the un-pruned tree.

```{r pruned tree}
# 2. pruned decision tree 
cv.tree.class <- cv.tree(tree.class, FUN = prune.misclass)
plot(cv.tree.class)
prune.tree.class <- prune.misclass(tree.class, best = 8)
plot(prune.tree.class); text(prune.tree.class, pretty = 0)
prune.tree.class.pred <- predict(prune.tree.class, tst, type = 'class')
prune.tree.class.train_err <- mean(trn$class != predict(prune.tree.class, type = 'class'))
prune.tree.class.test_err <- mean(tst$class != prune.tree.class.pred)
# prune.tree.class.test_err
```

The pruned tree produces a test error rate to the original tree with a misclassification error of `r prune.tree.class.test_err`.Using a Random Forest tree can often improve upon the error rate of a simple classification tree. While the test MSE was small in the classification tree, the team believed the error rate could be decreased using a more complex tree based method.

## Random Forest Classification
The random forest model is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees. Many bootstrap samples are created with replacement and then a classification tree is fitted on each bootstrap sample with a random subset of the predictors. Once a prediction has been made by the bootstrapped trees, the final classification is based on majority vote of the bootstrap predictions. Similarly, the prediction in the regression context is an average of the predictions of the bootstrapped trees.

The advantage for using an ensemble method over a regular classification or regression tree is that because there are many bootstrap samples being averaged together, there is less variance. This is similar to how the variance of the sample mean goes down as the sample size increases. Although it will probably increase our prediction accuracy, the random forest loses the interpretability that the basic trees have. Furthermore, the algorithm that is used to fit the random forest is very computationally expensive.

The first thing that should be done in building a boosted tree is cross validation. Here the model is built many times with the number of predictors included from one to all predictors. The test MSE is evaluated for each and the smallest value is selected. A plot of these is included below, the tree that produced the smallest error rate was the tree built with two variables, so that was the tree selected through cross validation.

```{r rf classification CV}
#----- classification -----
set.seed(702)
test.err <- double(11)
for (mtry in 1:11){
  fit <- randomForest(class ~ . - quality, trn, mtry = mtry)
  pred <- predict(fit, tst)
  test.err[mtry] <- mean(tst$class != pred)
}

# Plot cross validation results
matplot(1:mtry, test.err, pch = 1, type = 'b',
        xlab = "mtry", ylab = 'Mean Square Error')
points(which.min(test.err), test.err[which.min(test.err)], pch = 8, col = "#CC0000")
```

```{r rf class models}
rf.class <- randomForest(class ~ . - quality, trn, mtry = 2)
rf.class.pred <- predict(rf.class,tst)
rf.class.train_err <- mean(trn$class != predict(rf.class))
rf.class.test_err <- mean(tst$class != rf.class.pred)
# rf.class.train_err
# rf.class.test_err
```

Once the Random Forest tree with two variables was built, the final test MSE was calculated. This model produced a test MSE of `r round(rf.class.test_err, 3)`, which is significantly lower than the test MSE for the basic classification tree (`r round(prune.tree.class.test_err`)). The random forest was selected as the best method for predicting if a wine were to fall into the "high" or "low" class when it comes to wine quality.

# Regression Methods
## Multiple Linear Regression
NEED CODE FROM YUN
```{r mlr}
#### REGRESSION ####
lm.reg <- lm(quality ~ . - class, data = trn)

# Obtain RMSE and results for linear model
lm.reg.pred <- predict(lm.reg, newdata = tst)
lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
```
Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables. The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables. Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors. The formula for a multiple linear regression model with $p$ predictor variables is as follows:  
$$
\begin{aligned}
&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
& \quad \text{where for } i = n \text{ observations}, \\
& \quad y_i = \text{dependent variable (response)} \\
& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
& \quad \beta_0 = \text{y-intercept} \\
& \quad \beta_j = \text{slope coefficients for each variable} \\
& \quad \epsilon_i = \text{residuals (error term of model)} \\
\end{aligned}
$$

The multiple linear regression model is subject to the following assumptions about the data:

  * Linear relationship between response and predictors
  * No collinearity between predictor variables
  * $y_i$'s are iid
  * $\epsilon_i \sim N(0, \sigma^2)$

To assess these assumptions, we looked at various diagnostics plots:
```{r lm.diag, out.height = '3.5in'}
# Look at residual plot for nonlinearity
regres <- ggplot(data = data.frame(lm.reg$residuals, lm.reg$fitted.values), 
                 aes(x = lm.reg$residuals, y = lm.reg$fitted)) + 
            geom_point(position = 'jitter') + labs(x = "Residuals", y = "Fitted Values")

# Look at standardized residual plot for outliers
stdres <- ggplot(data = data.frame(x = 1:length(lm.reg$residuals), 
                                   y = scale(lm.reg$residuals)), 
                 aes(y = y, x = x)) + 
            geom_point() + labs(x = "Index", y = "Standardized Residuals")

# Look at a qqplot to assess normality
ggqq <- ggplot(data.frame(lm.reg$residuals), aes(sample = lm.reg$residuals)) + 
          stat_qq() + stat_qq_line(col = "blue", lty = 2) + 
          labs(x = "Theoretical", y = "Sample")

# Look at residuals vs leverage plot
reslev <- ggplot(lm.reg, aes(x = .hat, y = .stdresid)) + 
            geom_point(aes(size = .cooksd)) + 
            stat_smooth(method = "lm") + 
            labs( x = "Leverage", y = "Standardized Residuals") + 
            scale_size_continuous("Cook's D", range = c(1,5)) +
            theme_bw() + theme(legend.position = "right")

# Stitch all the graphs together
grid.arrange(regres, stdres, ggqq, reslev, nrow = 2)
```

~ talk about residual plots here ~

The biggest benefit to using a multiple linear regression model is its interpretability and ease of use. The coefficients can be computed efficiently and they are have clear and easy interpretations. It is also possible to make inference on any of the variables or linear combination of variables. The most signficiant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.

## Random Forest for Regression
As mentioned earlier in the classification section, Random Forest is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees. This model will be built exactly in the same way as the classification model, except the tree will not be predicting if the wine is a part of the good or bad class, but rather predicts the quality score associated with each wine. As before the model will first be built using cross validation to determine the number of variables that should be used in order to produce the smallest test MSE. Similar to the random forest classification tree, the cross validation selects the tree with 2 variables.

```{r rf regression}
#----- regression -----
# loop thru all the different values of mtry
test.err = double(11)
for (mtry in 1:11){
  fit<- randomForest(quality ~ . - class, trn, mtry = mtry)
  pred <- predict(fit, tst)
  test.err[mtry] <- sqrt(mean((tst$quality - pred)^2))
}

# Plot cross validation results
matplot(1:mtry, test.err, pch = 1, type = 'b', ylab = 'Mean Square Error', xlab = "mtry")
points(which.min(test.err), test.err[which.min(test.err)], pch = 8, col = "#CC0000")

# fit the random forest with the best results
rf.reg <- randomForest(quality ~ . - class, trn, mtry = 2)
rf.reg.pred <- predict(rf.reg, tst)
rf.reg.train_err <- sqrt(mean((trn$quality - predict(rf.reg))^2))
rf.reg.test_err <- sqrt(mean((tst$quality - rf.reg.pred)^2))
#rf.reg.train_err
#rf.reg.test_err
```  
The test MSE for the Regression Random Forest using 2 variables is 0.55. COMPARE THIS TO MLR WHEN WE GET YUN'S CODE.

# Results

```{r create.df}
# Create data frame output for regression models
df.reg <- rbind.data.frame(lm.reg.rmse, rf.reg.test_err) %>% round(4)
df.reg$method <- reg
colnames(df.reg) <- c(LETTERS[1:2])
df.reg <- df.reg %>% select("B", "A")
colnames(df.reg) <- c("Regression Method", "Root MSE")

# Create data frame output for classification models
df.class <- rbind.data.frame(prune.tree.class.test_err, rf.class.test_err) %>% round(4)
colnames(df.class) <- c("Misclassification Rate")
rownames(df.class) <- c("Classification Tree", "Random Forest")
```

```{r reg.out}
# Smash tables next to each other using kable
kable(df.reg, caption = "Regression Model Results", format = "latex") %>% 
  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
  kable_styling(latex_options = "HOLD_position")
```

```{r class.out}
kable(df.class, caption = "Classification Model Results", format = "latex") %>% 
  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
  kable_styling(latex_options = "HOLD_position")
```

Based on the output, the best model for both classification and regression was the random forest. This was not a surprise, since the ensemble methods tend to outperform regular methods. As an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance. We suspect that other ensemble methods, such as bagging or boosting, may also be able to yield higher prediction accuracy.

One of the main disadvantages to using ensemble techniques like the random forest is the fact that they tend to be very computationally intensive. The basic classification tree and multiple linear regression model took significantly less time to run than the random forest, but they were less accurate in their predictions. If we wanted to run an ensemble method on a very large data set with a lot of features, it would be a computational nightmare.

# Conclusion  
With the recent push for quantitative solutions in industries like agriculture, there has been an emphasis on using machine learning to solve a variety of problems requiring prediction and inference. In particular, the viticulture industry stands to gain a lot from such methods, since historically it has not interacted much with the world of analytics. In this study, we went through various machine learning techniques, assessing prediction accuracy in both classification and regression contexts.

Based on the results of our study, we were able to predict the high or low quality of a given wine with about 70-80% accuracy. Although this is a *good* result considering the non-information rate of 53%, it is by no means excellent. These results indicate that machine learning algorithms could be a good technique when trying to predict wine quality, but that there is room for improvement. With a more formal, industry expert-guided variable selection process, we believe that higher prediction accuracies can be achieved. Machine learning models could potentially be combined with other techniques to add value to winemakers.

## Limitations  
The multiple linear regression model is subject to certain distributional and model assumptions, which were not explicitly tested or checked. Since the objective of this project was to predict rather than make inference, we didn't explicitly check the model assumptions.

All of the wines in the dataset come from wineries in Portugal, and although they are from different wineries, we worry that the results may not be generalizable to wine production globally. Similarly, will these prediction results hold when it comes to white wine, rose, or sparkling wine? Given this variability over location and type of wine, what can we say about the winemaking industry as a whole?

Some of the wine regions have more strict standards for the designations of wine quality. For example, regions like Bordeaux and Champagne in France have some of the highest standards in the world for their wines, which are much different than somewhere like Greece or New Zealand. How accurate will are predictions be when there are different standards and how might we account for that variation?

## Further Research  
Since this study solely looks at the prediction accuracy of machine learning techniques for red wines, the natural extension of the analysis would to verify that the results hold constant for white wines as well. The University of California, Irvine Machine Learning database has data on white wines, so that would be a good place to start.

The characteristics of any given wine are often derived from different geological and/or climatological
conditions. For example, a wine that comes from a region close enough to the water to get a sea breeze may have a hint of saltiness in its aroma or taste. Similarly, a grape coming from a relatively young grove might give the wine different features than if the grove were hundreds of years old. It would be interesting to see how different weather conditions, terroirs, plant ages, and slopes affect the outcome of the wine in terms of quality.

# NOT NEEDED ???

I'm not sure we need these next 2 sections because most of the methods we are using do model selection. I don't have Yun's code, but I'm assuming he does some sort of variable selection, so this part seems extraneous.

## Cross Validation
Again, our focus is on the classification accuracy of machine learning methods. In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the models. We split the data into a training and test data set in order to later evaluate the model's prediction accuracy or root mean squared error (RMSE). For this project we used a 80/20 split, training the data on the 80% and testing the trained models on the withheld 20%. We will then repeat this process over 5 folds of the data, averaging the results.

```{r CV, eval = F}
# Specify the cross-validation method (5-fold CV)
trctrl <- trainControl(method = "cv", number = 5)
```

## Variable Selection
Since we have two different tasks that we are trying to complete (classification and regression), we went about selecting the variables for each task differently. For the classification task, we ran an initial random forest model on the entire dataset with selected predictors. We then looked at the variable importance of the predictors to decide which variables we would use in our final models. Within the context of machine learning, variable importance refers to how much a given model "uses" that variable to make accurate predictions. In other words, the more a model relies on a variable to make predictions, the more important it is for the model.

```{r class.select, out.height='4in', eval = F}
# Write preliminary model  
model.class.pre <- class ~ .

# Fit preliminary random forest on entire data set
rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
                             ntree = 500, mtry = 3, importance = TRUE)

# Create data frame of variable importance
df.class.pre <- rf.class.pre$importance %>% as.data.frame()
X <- rownames(df.class.pre)

# Sort the data frame and prep for plot
VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)

# Plot variable importance
vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
                ggtitle("Variable Importance")
vimp.class
```

We decided to choose the five variables with the most variable importance, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`. The fact that density proved to be an important variable for classification was relatively surprising. One would think that the wine quality would be determined based on its attributes that are detectable by the human senses. For this same reason, alcohol, acidity, and sulphates made sense in terms of variable importance. Sulphates are essential for the preservation of the wine, an overly- or underly-acidic wine may be considered poor, and the alcohol content has a connection to the fermentation process, where the must (aka the grape juice before it undergoes fermentation) actually becomes wine.

When looking at the problem in terms of a regression context, we used best subsets selection for feature selection. Using the `leaps` library and the `regsubsets()` function in R, we are able to compare all of the possible models from the given set of predictors. Through best subset selection, we can get the set of predictors for a model with any given amount of predictors. We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.

```{r reg.select, out.height='3.5in', eval = F}
# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
smre <- lm.sub %>% summary()

# Plot the Cp, BIC, R^2, RSS results in one plot
par(mfrow = c(2,2))
plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")

# Best subset selection
best.betas <- coef(lm.sub, id = 6) %>% round(2)
```

Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors). Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model. This was the model favored by BIC, which gives the greatest penalty for complexity in models. The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`. All of the variables from the classification selection appear here with the exception of density. In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses. The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.

```{r form, eval = F}
# Write formula for regression models
form.reg <- quality ~ volatile.acidity + chlorides + pH + 
            total.sulfur.dioxide + sulphates + alcohol

# Write formula for classification models
form.class <- class ~ alcohol + sulphates + density +
              volatile.acidity + total.sulfur.dioxide
```